---
title: "Calibration and limitations"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Calibration and limitations}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options:
  markdown:
    wrap: 72
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

**What this answers.** How the two conversion assumptions (words per line, points per minute) work, how to check predictions against observed completion times, and the limitations of the approach.

**What you need.** A `.qsf`; for validation, completion times for your respondents (ideally with their routes).

**What it cannot establish.** An implied points-per-minute rate is rough, trim-sensitive and specific to one survey. Burden scores do not predict dropout, satisficing or data quality.

This vignette covers two assumptions in the pipeline --
words-to-lines and points-to-minutes -- and the limitations of the
approach as a whole. For the scoring rules see `vignette("gfs-scoring")`.

```{r setup}
library(surveyBurden)
demo <- system.file("extdata", "demo_travel_survey.qsf", package = "surveyBurden")
```

## words_per_line

In the original GfS scheme, a score is provided for each additional line of text in the survey. However, a `.qsf` does not provide the number of additional lines of text, as this may vary with the survey mode. Instead, it returns the words per question. The package, therefore, assumes that the number of words that make an additional line (`words_per_line` = default 12). This is a practical proxy, that you can set if you wish, and means that assigned GfS+ points per descriptive-text in your survey should be read as approximate. This assumption only affects descriptive / instruction-text scores and every other question type is unaffected.

You can set the words per line in your survey with:

```{r wpl}
burden_report(demo, words_per_line = 8, quiet = TRUE)$burden$points[1:2]
```

## points_per_minute

To convert GfS+ points to minutes of survey time, the GfS rule of thumb is 12 points per minute. As a rule of thumb, do not assume that it predicts the completion time, or applies to online surveys. You can either override it by setting `points_per_minute` on a weights object:

```{r ppm}
w <- gfs_scheme()
w$points_per_minute <- 15
burden_report(demo, scheme = w, quiet = TRUE)$burden[, c("statistic", "minutes")]
```

Alternatively, if you are calculating burden ex post then you can use the completion-time data of your survey to calculate `validate_times()` and use the rate it
returns.

### Validating against completion times

`validate_times()` predicts each respondent's burden from their route
(via `respondent_burden()`) and compares it with their completion time.
Prepare the data first; the function does little of this for you:

1. **Keep completed responses only.** `trim` filters on duration alone
   and does not detect non-finishers, so filter on `Finished` yourself
   (see `vignette("realised-burden")` for its encodings).
2. **Name the duration column** `completion_seconds` or
   `completion_mins`. Qualtrics exports it as `Duration (in seconds)`,
   which is not recognised. If you read a raw CSV, drop the two header
   rows first, or the column is read as text.
3. **Add route columns** (`loop_*`, `visit_*`) if you have them -- see
   `vignette("paths-and-display-logic")`. Without them every respondent
   is predicted at `loop_typical`.
4. **Choose the trim.** The default keeps 3 to 180 minutes, to drop
   speeders and people who left the survey open. Report how many rows it
   removed (`n` against your row count); the implied rate is sensitive to
   it.

A simulated example:

```{r validate}
set.seed(2)
n <- 120
paradata <- data.frame(
  ResponseId = paste0("R_", seq_len(n)),
  Finished   = rbinom(n, 1, 0.9),                    # 1 = completed
  `Duration (in seconds)` = round(rlnorm(n, log(600), 0.5)),
  loop_BL6   = rbinom(n, 5, 0.4),
  loop_BL8   = rbinom(n, 3, 0.5),
  check.names = FALSE
)

obs <- paradata[paradata$Finished == 1, ]            # step 1
obs$completion_seconds <- obs$`Duration (in seconds)` # step 2

vt <- validate_times(demo, obs, trim = c(3, 180))    # step 4
c(rows = nrow(paradata), finished = nrow(obs), kept_after_trim = vt$n)
round(c(ratio = vt$ratio, cor = vt$cor,
        implied_ppm = vt$implied_points_per_minute), 2)
```

Reading the output:

- `ratio` is the predicted median over the observed median, in minutes.
- `cor` is the correlation of predicted points with observed minutes.
  Route-level predictions cannot see within-path variation, so expect it
  to be small on real data.
- `implied_points_per_minute` is `60 / b`, where `b` is the slope, in
  seconds per point, of a through-origin fit of observed seconds on
  predicted points. It is a rough, trim-sensitive rate for this survey,
  not a calibrated constant.

To use that rate:

```{r validate-use}
w <- gfs_scheme()
w$points_per_minute <- vt$implied_points_per_minute
burden_report(demo, scheme = w, quiet = TRUE)$burden[, c("statistic", "minutes")]
```

## Limitations

- The Qualtrics-to-GfS+ mapping is sometimes inferential. Where a question type
  has no exact match in Table 1, the package applies a documented rule
  and flags the question `inferred`.
- The original GfS scheme predates modern web survey interfaces. It was
  calibrated on earlier survey practice.
- Rendered text length cannot be recovered from a `.qsf`. The file stores
  words, not an on-screen line count; `words_per_line` is a configurable
  proxy for that missing count, not a measured constant.
- Display-logic reconstruction depends on what the `.qsf` encodes. Logic
  controlled outside Qualtrics, or unusual constructs, may need manual
  interpretation.
- The structural profile is not a respondent distribution. Without route
  data, each complete path gets equal total weight and, within a path, each
  modelled display-logic state and loop count gets equal weight. "Median"
  means the median of that model-weighted distribution, not across
  respondents.
- Structural paths are the block sequences the flow allows. Branch
  conditions are not evaluated or checked against each other, so the
  minimum or maximum may not be reachable by a real respondent.
- Loop & Merge can make the path space large. The package unrolls to the
  explicit cap and treats the iteration count as uniform.
- The points-to-minutes conversion is the original GfS rule of thumb,
  roughly 12 points per minute.
- Burden scores are not predictions of completion time, dropout, or
  satisficing. Those are empirical questions about respondent behaviour,
  outside the scope of this package.

## Next

- `vignette("reading-the-report")` -- reading the full report.
- `vignette("realised-burden")` -- answer-based burden after fielding.

## References

- Heimgartner, D. and Axhausen, K. W. (2024). Predicting response rates
  once again. *Findings*.
  <https://findingspress.org/article/125481-predicting-response-rates-once-again>
- Yan, T. and Tourangeau, R. (2008). Fast times and easy questions: the
  effects of age, experience, and question complexity on web survey
  response times. *Applied Cognitive Psychology*.
