---
title: "Calibration and limitations"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Calibration and limitations}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options:
  markdown:
    wrap: 72
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

This vignette covers two assumptions in the pipeline --
words-to-lines and points-to-minutes -- and the limitations of the
approach as a whole. For the scoring rules see `vignette("gfs-scoring")`.

```{r setup}
library(surveyBurden)
demo <- system.file("extdata", "demo_travel_survey.qsf", package = "surveyBurden")
```

## words_per_line

In the original GfS papers, they provide a score for each additional line of text in the survey. However, a `.qsf` does not provide the number of additional lines of text, as this may vary with the survey mode. Instead, it returns the words per question. The package, therefore, assumes that the number of words that make an additional line (`words_per_line` = default 12). This is a practical proxy, that you can set if you wish, and means that assigned GfS points per descriptive-text in your survey should be read as approximate. This assumption only affects descriptive / instruction-text scores and every other question type is unaffected.

You can set the words per line in your survey with:

```{r wpl}
burden_report(demo, words_per_line = 8, quiet = TRUE)$burden$points[1:2]
```

## points_per_minute

To convert GfS points to minutes of survey time, they have a rule of thumb of 12 points per minute. As a rule of thumb, do not assume that it predicts the completion time, or applies to online surveys. You can either override it by setting `points_per_minute` on a weights object:

```{r ppm}
w <- gfs_scheme()
w$points_per_minute <- 15
burden_report(demo, scheme = w, quiet = TRUE)$burden[, c("statistic", "minutes")]
```

Alternatively, if you are calculating burden ex post then you can use the completion-time data of your survey to calculate `validate_times()` and use the rate it
returns:

```{r validate, eval = FALSE}
w <- gfs_scheme()
w$points_per_minute <- validate_times(demo, "times.csv")$implied_points_per_minute
burden_report(demo, scheme = w)
```

## Limitations

- The Qualtrics-to-GfS mapping is sometimes inferential. Where a question type
  has no exact match in Table 1, the package applies a documented rule
  and flags the question `inferred`.
- The GfS scheme predates modern web survey interfaces. It was calibrated
  on earlier survey practice.
- Rendered text length cannot be recovered from a `.qsf`. The file stores
  words, not an on-screen line count; `words_per_line` is a configurable
  proxy for that missing count, not a measured constant.
- Display-logic reconstruction depends on what the `.qsf` encodes. Logic
  controlled outside Qualtrics, or unusual constructs, may need manual
  interpretation.
- Ex ante structural enumeration of all the paths through your survey does not provide respondent probabilities. Without route
  data every feasible combination is counted once, and "median" means the
  median across combinations, not across respondents.
- Loop & Merge can make the path space large. The package unrolls to the
  explicit cap and treats the iteration count as uniform.
- The points-to-minutes conversion is the published GfS rule of thumb,
  roughly 12 points per minute.
- Burden scores are not predictions of completion time, dropout, or
  satisficing. Those are empirical questions about respondent behaviour,
  outside the scope of this package.

## References

- Heimgartner, D. and Axhausen, K. W. (2024). Predicting response rates
  once again. *Findings*.
  <https://findingspress.org/article/125481-predicting-response-rates-once-again>
- Yan, T. and Tourangeau, R. (2008). Fast times and easy questions: the
  effects of age, experience, and question complexity on web survey
  response times. *Applied Cognitive Psychology*.
