---
title: "Extensions: scoring surveys without a QSF"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Extensions: scoring surveys without a QSF}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options:
  markdown:
    wrap: 72
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

**What this answers.** What you can do with a survey not built in Qualtrics, and what full support for another platform would take.

**What you need.** A question catalogue describing your survey (schema below).

**What it cannot establish.** Without routing and display-logic information, a catalogue gives per-question scores and a single-pass total, not a burden range or profile.

```{r setup}
library(surveyBurden)
```

surveyBurden reads Qualtrics `.qsf` files, but many surveys are built on
other platforms (LimeSurvey, SurveyEngine, REDCap, and so on). This
vignette documents where the Qualtrics-specific code ends, what you can
already do with a survey from another platform, and what supporting a new
platform fully would take.

In short:

| Status | What | How |
|---|---|---|
| **Working now** | Per-question GfS+ scores and a survey total for any survey you can describe as a list of questions | `validate_catalogue()` then `score_burden()` |
| **Requires an adapter** | A floor/ceiling burden band per path | build a paths table yourself, then `path_burden(NULL, paths = , scored = )` |
| **Future interface work** | Display-logic profiles, `burden_report()`, respondent predictions, realised burden, calculation certainty | these currently need a Qualtrics `qsf_raw` object |

## Three-layer architecture

The package has three layers:

```
Layer 1:  Parser          (platform-specific)
          read_qsf() / fetch_qsf() → parse_qsf()
                                  ↓
          Question catalogue  ← the intermediate representation
                                  ↓
Layer 2:  Path resolver     (partially platform-specific)
          resolve_flow() → resolve_paths()
                                  ↓
Layer 3:  Scorer + reporter (platform-agnostic)
          score_burden() → path_burden() → burden_report()
```

**Layer 1** reads a Qualtrics survey and produces a
standardised *question catalogue*: a tibble with one row per live
question, in flow order, with the structural fields the scorer needs.
This is the intermediate representation.

**Layer 2** resolves the survey's routing logic (branches, block
randomisers, early exits) and display logic (conditional questions) into
the structural paths the flow allows. This layer reads the QSF's
`SurveyFlow`, block and `DisplayLogic` elements directly, so it is
Qualtrics-specific.

**Layer 3** applies the GfS+ points to the catalogue, computes per-path
burden bands and profiles, and assembles the report. Only part of it is
independent of the QSF: `score_burden()` needs just the catalogue, and
`path_burden()` needs just the catalogue and a paths table. The profile
engine behind `path_burden_profile()` and `burden_report()` still reads
block definitions and display-logic trees from the `qsf_raw` object.

## The question catalogue

The catalogue is a tibble produced by `parse_qsf()`, one row per live
question. Any parser that produces this schema can use `score_burden()`
directly. The functions further down the pipeline need more than the
catalogue; see [What a new parser must provide](#what-a-new-parser-must-provide).

### Required columns

| Column | Type | Description |
|--------|------|-------------|
| `question_id` | character | Unique identifier for each question |
| `std_type` | character | Standardised question type (see below) |

### Structural columns (used by the scorer)

| Column | Type | Description |
|--------|------|-------------|
| `n_options` | integer | Number of response options (`NA` if unknown) |
| `n_rows` | integer | Number of matrix rows (`NA` for non-matrix) |
| `n_cols` | integer | Number of matrix columns (`NA` for non-matrix) |
| `text_words` | integer | Word count of the question stem |
| `is_dropdown` | logical | Single-choice rendered as a dropdown |
| `is_multiline` | logical | Open-text field with multi-line entry |
| `is_multi_answer` | logical | Matrix where each cell is a checkbox |
| `is_hidden` | logical | Hidden from the respondent |
| `options_numeric` | logical | All option labels are numbers |
| `label_text` | character | Lowercased, joined response labels |
| `question_text` | character | Plain-text question stem (for cue detection) |
| `has_display_logic` | logical | Whether the question is conditionally shown |
| `has_validation` | logical | Whether a response is forced |
| `in_loop` | logical | Whether the question is inside a loop block |
| `loop_max` | integer | Maximum loop iterations (`NA` if not in loop) |
| `flag` | character | Confidence: `"auto"`, `"inferred"`, `"manual"`, `"unknown"` |

### Provenance columns (not used by the scorer)

| Column | Type | Description |
|--------|------|-------------|
| `qualtrics_type` | character | Original platform question type |
| `selector` | character | Platform-specific rendering selector |
| `subselector` | character | Platform-specific sub-selector |
| `flow_order` | integer | Position in the survey flow |
| `block_id` | character | Block identifier |
| `block_name` | character | Block description |

The `qualtrics_type`, `selector`, and `subselector` columns are carried
for debugging and provenance. The scorer does not read them; it uses
`std_type` and the boolean columns (`is_dropdown`, `is_multiline`,
`is_multi_answer`) instead.

### Standardised question types

The `std_type` column uses the following vocabulary:

| `std_type` | Description | GfS+ scoring |
|------------|-------------|-------------|
| `single_choice` | One answer from a list (radio, dropdown, NPS) | By option count |
| `multi_choice` | Multiple answers from a list (checkboxes) | By option count |
| `matrix` | Grid/table: rows x columns | Per row, by column count |
| `open_text` | Free text (single-line or essay) | Short vs essay |
| `slider` | Continuous input (visual analogue) | By label cues |
| `ranking` | Drag-and-drop or number ranking | Fixed |
| `constant_sum` | Allocate a total across categories | As numeric answer |
| `descriptive` | Instruction or transition text (no input) | By word count |
| `timing` | Page-timing metadata (not shown) | 0 points |
| `meta` | Browser/metadata capture (not shown) | 0 points |
| `captcha` | Bot-detection challenge | Fixed |

A new parser maps its platform's question types into this vocabulary.
The mapping need not be one-to-one: several Qualtrics types map to
`single_choice`, for example.

## Validating a hand-built catalogue

`validate_catalogue()` checks that a data frame has the columns the
scorer needs, fills missing optional columns with safe defaults, and
coerces the result to a tibble:

```{r}
cat <- data.frame(
  question_id = c("Q1", "Q2", "Q3", "Q4", "Q5"),
  std_type    = c("single_choice", "multi_choice", "matrix",
                  "open_text", "descriptive"),
  n_options   = c(5L, 8L, NA, NA, NA),
  n_rows      = c(NA, NA, 4L, NA, NA),
  n_cols      = c(NA, NA, 5L, NA, NA),
  text_words  = c(8L, 12L, 6L, 10L, 25L),
  is_multiline = c(FALSE, FALSE, FALSE, TRUE, FALSE),
  stringsAsFactors = FALSE
)
validated <- validate_catalogue(cat)
names(validated)
```

## Worked example: scoring without a QSF

This example constructs a five-question catalogue by hand and runs it
through `score_burden()`. No QSF file is involved.

```{r}
catalogue <- validate_catalogue(data.frame(
  question_id = c("Q1", "Q2", "Q3", "Q4", "Q5"),
  std_type    = c("single_choice", "multi_choice", "matrix",
                  "open_text", "descriptive"),
  n_options   = c(5L, 8L, NA, NA, NA),
  n_rows      = c(NA, NA, 4L, NA, NA),
  n_cols      = c(NA, NA, 5L, NA, NA),
  text_words  = c(8L, 12L, 6L, 10L, 25L),
  is_multiline = c(FALSE, FALSE, FALSE, TRUE, FALSE),
  question_text = c(
    "How often do you cycle?",
    "Which of these apply?",
    "Rate each mode (1-5)",
    "Describe your typical commute",
    "The next section asks about your household"
  ),
  stringsAsFactors = FALSE
))

scored <- score_burden(catalogue)
scored[, c("question_id", "std_type", "gfs_points",
           "score_flag", "score_basis")]
```

The function score_burden operates entirely on the question catalogue so that it does not know or care
that these questions came from a hand-built data frame rather than a
Qualtrics `.qsf`.

### Survey-level burden

If using the validate_catalogue approach to calculate the burden of a non-Qualtrics survey, you could sum the per-question scores for the survey-level total:

```{r}
total_points <- sum(scored$gfs_points, na.rm = TRUE)
total_minutes <- total_points / gfs_scheme()$points_per_minute
cat(sprintf("Total: %.0f GfS+ points, ~%.0f minutes\n",
            total_points, total_minutes))
```

This is the burden of one path that shows every question once. For a
survey with conditional questions it is an upper bound on a single pass
-- the burden a respondent would face if every question were shown -- but
not if questions repeat in a loop: multiply looped questions by their
iteration count yourself.

### What this gives you and what it does not

A hand-built catalogue scored with `score_burden()` gives you:

-   Per-question GfS+ burden points, using the same scoring rules as
    `burden_report()`.
-   A survey-level total (the `sum()` above).
-   A per-question confidence flag (`score_flag`) saying how the score
    was reached.

It does **not** give you:

-   A burden *range* across branching paths -- that requires a paths
    table (see the adapter below).
-   A display-logic profile -- the structural min/median/max within a
    path. That currently requires a Qualtrics survey.
-   A population-weighted or realised burden -- `respondent_burden()` and
    `realised_burden()` currently require a Qualtrics survey as well as
    response data.

The total from a hand-built catalogue is a defensible measure of
instrument burden for any survey that can be expressed as a list of
questions with their types and option counts. It applies the same
published GfS+ point scheme (Heimgartner and Axhausen, 2024) as the
full Qualtrics pipeline.

## What a new parser must provide

### Working now: a question catalogue

A data frame matching the schema above, passed through
`validate_catalogue()`. This is all `score_burden()` needs: per-question
scores and a survey total work immediately, as in the worked example.

### Requires an adapter: a paths table

`path_burden()` gives a floor/ceiling band per path without a QSF if you
pass it the scored catalogue and a paths table with one row per path:
`path_id`, `terminates_early`, `block_ids` (list), `q_always` (list of
question ids always shown on that path), `q_maybe` (list of question ids
that may be shown) and `n_gates`.

```{r adapter}
paths <- tibble::tibble(
  path_id          = 1:2,
  terminates_early = c(FALSE, FALSE),
  block_ids        = list("B1", c("B1", "B2")),
  q_always         = list(c("Q1", "Q2", "Q3"), c("Q1", "Q2", "Q3", "Q4")),
  q_maybe          = list(character(0), "Q5"),
  n_gates          = c(0L, 1L)
)
path_burden(NULL, paths = paths, scored = scored)[,
  c("path_id", "gfs_floor", "gfs_ceiling")]
```

Building that table is the adapter's job: enumerating the platform's
paths and deciding, for each, which questions are always or maybe shown.
Loop multiplication in the ceiling uses the catalogue's `loop_max`.

### Future interface work: routing and display logic

Everything beyond the band -- `path_burden_profile()`, `burden_report()`,
`respondent_burden()`, `calculation_certainty()` and `realised_burden()`
-- currently takes a `qsf_raw` object and reads its `SurveyFlow`, block
and `DisplayLogic` elements directly. Supporting another platform there
needs one of:

- **Translation**: convert the platform's survey into a `qsf_raw`-shaped
  object -- flow nodes (`Standard`, `Branch`, `EndSurvey`,
  `BlockRandomizer`, `Group`, `EmbeddedData`), block definitions, and
  question payloads with Qualtrics-format `DisplayLogic` trees. This
  works with the current code but is a substantial mapping.
- **A platform-neutral interface** (not yet built): letting these
  functions accept a catalogue, a paths table and parsed display-logic
  predicates in place of `qsf_raw`.

Setting `has_display_logic = FALSE` for every question does **not** skip
the display-logic profile: it tells the package the question is always
shown, so conditional questions count on every path and burden is
overstated. It removes conditionality from the model rather than
modelling it.

## Where platform-specific code lives

| Function | Layer | Platform-specific? |
|----------|-------|--------------------|
| `read_qsf()` / `fetch_qsf()` | 1 | Yes -- reads QSF format |
| `classify_question()` | 1 | Yes -- maps Qualtrics type codes |
| `parse_qsf()` | 1 | Yes -- orchestrates QSF parsing |
| `resolve_live_blocks()` | 1-2 | Yes -- reads QSF block definitions |
| `resolve_flow()` | 2 | Yes -- walks QSF `SurveyFlow` nodes |
| `parse_display_logic()` | 2 | Yes -- reads QSF `DisplayLogic` trees |
| `validate_catalogue()` | boundary | No -- validates the intermediate representation |
| `score_burden()` | 3 | No -- operates on the catalogue |
| `gfs_scheme()` | 3 | No -- scoring parameters |
| `classify_reachability()` | 2-3 | Partly -- catalogue fields, but ruling questions out needs Qualtrics-format `DisplayLogic` trees |
| `path_burden()` | 3 | No, given a paths table and the scored catalogue |
| `path_burden_profile()` | 3 | Yes, at present -- its engine reads blocks and display logic from `qsf_raw` |
| `respondent_burden()`, `calculation_certainty()`, `realised_burden()` | 3 | Yes, at present -- take `qsf_raw` |
| `burden_report()` | 3 | Orchestrator -- takes `qsf_raw` and threads it through Layers 1-2 |

The boundary between platform-specific and platform-agnostic code
currently sits at the question catalogue for scoring, and at a paths
table for the floor/ceiling band. Everything else still reads the QSF.

## Design rationale

surveyBurden is a Qualtrics-first tool and does not claim to be
platform-agnostic. The scoring rules are already separate from the
parser, so scoring another platform's survey needs only a catalogue.
Full support would additionally need:

1. A parser that produces the question catalogue (the main work).
2. Routing: either translation into Qualtrics flow nodes, or the
   platform-neutral interface above.
3. Display logic, likewise, if the platform has per-question visibility
   conditions.

For a platform with no branching, loops or conditional questions, the
catalogue and `score_burden()` already give the full answer.

## Next

- `vignette("gfs-scoring")` -- the scoring rules the catalogue feeds.
- `?validate_catalogue` -- the catalogue checks and defaults.

