---
title: "Extensions: scoring surveys without a QSF"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Extensions: scoring surveys without a QSF}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options:
  markdown:
    wrap: 72
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

```{r setup}
library(surveyBurden)
```

surveyBurden reads Qualtrics `.qsf` files. However, Qualtrics is closed-source and not universal. Instead, a better scenario would be for surveyBurden to separate the platform-specific parser from the
platform-agnostic scoring and analysis layers. In this vignette, we document
that boundary so that someone who wanted to add support for another
platform (LimeSurvey, SurveyEngine, REDCap, etc.) could write a new
parser without touching the scoring or path-analysis code.

## Three-layer architecture

The package has three layers:

```
Layer 1:  Parser          (platform-specific)
          read_qsf() / fetch_qsf() → parse_qsf()
                                  ↓
          Question catalogue  ← the intermediate representation
                                  ↓
Layer 2:  Path resolver     (partially platform-specific)
          resolve_flow() → resolve_paths()
                                  ↓
Layer 3:  Scorer + reporter (platform-agnostic)
          score_burden() → path_burden() → burden_report()
```

**Layer 1** reads a Qualtrics survey and produces a
standardised *question catalogue*: a tibble with one row per live
question, in flow order, with the structural fields the scorer needs.
This is the intermediate representation.

**Layer 2** resolves the survey's routing logic (branches, block
randomisers, early exits) and display logic (conditional questions) into
a set of feasible respondent paths. This layer currently reads the QSF's
`SurveyFlow` element directly and is, therefore, partially
platform-specific. Non-Qualtrics platforms would need to provide an equivalent
flow structure or an adapter (see *What a new parser must provide*
below).

**Layer 3** applies the GfS+ points to the catalogue, computes
per-path burden bands, and assembles the report. This layer operates
entirely on the question catalogue and the resolved path structure. It
never touches the QSF, so it is feasibly platform-agnostic.

## The question catalogue

The catalogue is a tibble produced by `parse_qsf()`, one row per live
question. Any parser that produces this schema can use `score_burden()`
and downstream functions unchanged.

### Required columns

| Column | Type | Description |
|--------|------|-------------|
| `question_id` | character | Unique identifier for each question |
| `std_type` | character | Standardised question type (see below) |

### Structural columns (used by the scorer)

| Column | Type | Description |
|--------|------|-------------|
| `n_options` | integer | Number of response options (`NA` if unknown) |
| `n_rows` | integer | Number of matrix rows (`NA` for non-matrix) |
| `n_cols` | integer | Number of matrix columns (`NA` for non-matrix) |
| `text_words` | integer | Word count of the question stem |
| `is_dropdown` | logical | Single-choice rendered as a dropdown |
| `is_multiline` | logical | Open-text field with multi-line entry |
| `is_multi_answer` | logical | Matrix where each cell is a checkbox |
| `is_hidden` | logical | Hidden from the respondent |
| `options_numeric` | logical | All option labels are numbers |
| `label_text` | character | Lowercased, joined response labels |
| `question_text` | character | Plain-text question stem (for cue detection) |
| `has_display_logic` | logical | Whether the question is conditionally shown |
| `has_validation` | logical | Whether a response is forced |
| `in_loop` | logical | Whether the question is inside a loop block |
| `loop_max` | integer | Maximum loop iterations (`NA` if not in loop) |
| `flag` | character | Confidence: `"auto"`, `"inferred"`, `"manual"`, `"unknown"` |

### Provenance columns (not used by the scorer)

| Column | Type | Description |
|--------|------|-------------|
| `qualtrics_type` | character | Original platform question type |
| `selector` | character | Platform-specific rendering selector |
| `subselector` | character | Platform-specific sub-selector |
| `flow_order` | integer | Position in the survey flow |
| `block_id` | character | Block identifier |
| `block_name` | character | Block description |

The `qualtrics_type`, `selector`, and `subselector` columns are carried
for debugging and provenance. The scorer does not read them; it uses
`std_type` and the boolean columns (`is_dropdown`, `is_multiline`,
`is_multi_answer`) instead.

### Standardised question types

The `std_type` column uses the following vocabulary:

| `std_type` | Description | GfS+ scoring |
|------------|-------------|-------------|
| `single_choice` | One answer from a list (radio, dropdown, NPS) | By option count |
| `multi_choice` | Multiple answers from a list (checkboxes) | By option count |
| `matrix` | Grid/table: rows x columns | Per row, by column count |
| `open_text` | Free text (single-line or essay) | Short vs essay |
| `slider` | Continuous input (visual analogue) | By label cues |
| `ranking` | Drag-and-drop or number ranking | Fixed |
| `constant_sum` | Allocate a total across categories | As numeric answer |
| `descriptive` | Instruction or transition text (no input) | By word count |
| `timing` | Page-timing metadata (not shown) | 0 points |
| `meta` | Browser/metadata capture (not shown) | 0 points |
| `captcha` | Bot-detection challenge | Fixed |

A new parser maps its platform's question types into this vocabulary.
The mapping need not be one-to-one: several Qualtrics types map to
`single_choice`, for example.

## Validating a hand-built catalogue

`validate_catalogue()` checks that a data frame has the columns the
scorer needs, fills missing optional columns with safe defaults, and
coerces the result to a tibble:

```{r}
cat <- data.frame(
  question_id = c("Q1", "Q2", "Q3", "Q4", "Q5"),
  std_type    = c("single_choice", "multi_choice", "matrix",
                  "open_text", "descriptive"),
  n_options   = c(5L, 8L, NA, NA, NA),
  n_rows      = c(NA, NA, 4L, NA, NA),
  n_cols      = c(NA, NA, 5L, NA, NA),
  text_words  = c(8L, 12L, 6L, 10L, 25L),
  is_multiline = c(FALSE, FALSE, FALSE, TRUE, FALSE),
  stringsAsFactors = FALSE
)
validated <- validate_catalogue(cat)
names(validated)
```

## Worked example: scoring without a QSF

This example constructs a five-question catalogue by hand and runs it
through `score_burden()`. No QSF file is involved.

```{r}
catalogue <- validate_catalogue(data.frame(
  question_id = c("Q1", "Q2", "Q3", "Q4", "Q5"),
  std_type    = c("single_choice", "multi_choice", "matrix",
                  "open_text", "descriptive"),
  n_options   = c(5L, 8L, NA, NA, NA),
  n_rows      = c(NA, NA, 4L, NA, NA),
  n_cols      = c(NA, NA, 5L, NA, NA),
  text_words  = c(8L, 12L, 6L, 10L, 25L),
  is_multiline = c(FALSE, FALSE, FALSE, TRUE, FALSE),
  question_text = c(
    "How often do you cycle?",
    "Which of these apply?",
    "Rate each mode (1-5)",
    "Describe your typical commute",
    "The next section asks about your household"
  ),
  stringsAsFactors = FALSE
))

scored <- score_burden(catalogue)
scored[, c("question_id", "std_type", "gfs_points",
           "score_flag", "score_basis")]
```

The function score_burden operates entirely on the question catalogue so that it does not know or care
that these questions came from a hand-built data frame rather than a
Qualtrics `.qsf`.

### Survey-level burden

If using the validate_catalogue approach to calculate the burden of a non-Qualtrics survey, you could sum the per-question scores for the survey-level total:

```{r}
total_points <- sum(scored$gfs_points, na.rm = TRUE)
total_minutes <- total_points / gfs_scheme()$points_per_minute
cat(sprintf("Total: %.0f GfS+ points, ~%.0f minutes\n",
            total_points, total_minutes))
```

This provides a headline estimate of the burden across one single path through the survey i.e., every question is counted once. For a
survey with conditional questions, it is an upper bound -- the burden a
respondent would face if every question were shown. Future work will add support for display logic.

### What this gives you and what it does not

A hand-built catalogue scored with `score_burden()` gives you:

-   Per-question GfS+ burden points, using the same scoring rules as
    `burden_report()`.
-   A survey-level total (the `sum()` above).
-   A per-question confidence flag (`score_flag`) saying how the score
    was reached.

It does **not** give you:

-   A burden *range* across branching paths -- that requires flow
    resolution, which needs a QSF or an equivalent flow structure.
-   A display-logic profile -- the structural min/median/max within a
    path. That requires the parsed display-logic predicates.
-   A population-weighted burden -- that requires response data
    (`realised_burden()`).

The total from a hand-built catalogue is a defensible measure of
instrument burden for any survey that can be expressed as a list of
questions with their types and option counts. It applies the same
published GfS+ point scheme (Heimgartner and Axhausen, 2024) as the
full Qualtrics pipeline.

## What a new parser must provide

To support a new survey platform, you need to provide:

### 1. A question catalogue (required)

A data frame matching the schema above, passed through
`validate_catalogue()`. This is sufficient for `score_burden()` --
per-question burden scoring works immediately.

### 2. A flow structure (for path analysis)

`resolve_flow()` currently walks the QSF's `SurveyFlow` node tree,
which uses Qualtrics-specific node types (`Standard`, `Branch`,
`EndSurvey`, `BlockRandomizer`, `Group`, `EmbeddedData`). The function
returns a tibble of `(path_id, block_ids, terminates_early, decisions)`.

There are two options for extending surveyBurden to a new platform:

- **1) Adoption**: translate the platform's routing model into
  the same node-list format `enumerate_paths()` consumes. This
  requires mapping the platform's branching constructs to `Branch`
  nodes, its randomisation to `BlockRandomizer`, etc.
- **2) Replacement**: write a platform-specific path enumerator
  that returns the same output schema (a tibble with `path_id`,
  `block_ids`, `terminates_early`, `decisions` columns).

Either way, `path_burden()` and `path_burden_profile()` work unchanged
once they receive the resolved paths.

### 3. Display logic (for the burden profile)

`parse_display_logic()` reads the QSF's `DisplayLogic` tree format. The
display-logic layer is the most tightly coupled to Qualtrics. A new
platform would need to produce an equivalent parsed structure: for each
conditional question, the set of gate variables it reads and a predicate
function that evaluates the logic against an assignment of those
variables. See `?classify_reachability` for how the downstream code
consumes display-logic information.

For a first implementation, setting `has_display_logic = FALSE` for all
questions is a valid starting point. The scorer and per-path burden
still work; only the within-path display-logic profile
(`path_burden_profile()`) is skipped.

## Where platform-specific code lives

| Function | Layer | Platform-specific? |
|----------|-------|--------------------|
| `read_qsf()` / `fetch_qsf()` | 1 | Yes -- reads QSF format |
| `classify_question()` | 1 | Yes -- maps Qualtrics type codes |
| `parse_qsf()` | 1 | Yes -- orchestrates QSF parsing |
| `resolve_live_blocks()` | 1-2 | Yes -- reads QSF block definitions |
| `resolve_flow()` | 2 | Yes -- walks QSF `SurveyFlow` nodes |
| `parse_display_logic()` | 2 | Yes -- reads QSF `DisplayLogic` trees |
| `validate_catalogue()` | boundary | No -- validates the intermediate representation |
| `score_burden()` | 3 | No -- operates on the catalogue |
| `gfs_scheme()` | 3 | No -- scoring parameters |
| `classify_reachability()` | 3 | No -- operates on catalogue fields |
| `path_burden()` | 3 | No -- operates on scored catalogue + paths |
| `path_burden_profile()` | 3 | No -- operates on the burden engine |
| `burden_report()` | 3 | Orchestrator -- takes `qsf_raw` but threads it through Layers 1-2 |

The boundary between platform-specific and platform-agnostic code sits
at the question catalogue. Everything below `score_burden()` is
reusable; everything above it (and the flow/display-logic resolvers) is
Qualtrics-specific.

## Design rationale

For now, this a Qualtrics-first tool that does not claim to be platform-agnostic. However, the three-layer design means that adding a new platform does
not require touching the scoring rules, the path enumeration algebra, or the reporting logic. What it does require is:

1. A parser that produces the question catalogue (the main work).
2. A flow enumerator if the platform has conditional routing.
3. A display-logic parser if the platform has per-question visibility
   conditions.

For platforms with simpler routing than Qualtrics (no branching, no
block randomisers), steps 2 and 3 may be trivial or unnecessary.
