---
title: "The GfS scoring method"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{The GfS scoring method}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options:
  markdown:
    wrap: 72
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

This article covers how one Qualtrics question becomes a GfS point score.
For how those scores are combined across the survey's routes see
`vignette("paths-and-display-logic")`; for reading the finished report
see `vignette("reading-the-report")`.

```{r setup}
library(surveyBurden)
demo <- system.file("extdata", "demo_travel_survey.qsf", package = "surveyBurden")
qsf <- read_qsf(demo)
catalogue <- parse_qsf(qsf)
```

surveyBurden implements the published GfS / Axhausen scheme (Heimgartner
and Axhausen 2024, Table 1); it does not invent a new burden scale. The
point weights are transcribed verbatim into `gfs_weights()`.

```{r weights}
w <- gfs_weights()
w$yes_no
w$rating_small
w$rating_large
w$points_per_minute
```

A few of the Table 1 weights, in points per response action:

| response action | GfS points |
|---|---|
| yes / no question | 1 |
| rating scale, up to 5 options | 2 |
| rating scale, more than 5 options | 3 |
| first answer to an open question | 6 |
| two-alternative stated-choice task | 2 |

The full table is in Heimgartner and Axhausen (2024); it is not
reproduced here.

## Classification first

`classify_question()` maps each Qualtrics widget to a standard question
type. The Qualtrics type alone is not enough: one `MC` widget can be a
yes/no item, a rating scale, a dropdown, or a multi-select, depending on
its selector and options.

```{r classify}
table(catalogue$std_type)
```

## Three layers

`score_burden()` then works in three layers:

1. what Qualtrics says the widget is (for example `MC` with a dropdown
   selector);
2. what response action that implies (choose one value from an ordered
   list);
3. which Table 1 category best represents that action (a rating).

Where step 3 is fixed by the structure -- a two-option choice is a closed
yes/no item whatever it asks -- the mapping is deterministic and the
score is marked `auto`. Where the `.qsf` does not carry what the scheme
needs, the package applies an **explicit, documented inference rule** and
marks the item `inferred`. Every scored question carries a `score_basis`
string recording the rule applied.

```{r scored}
scored <- score_burden(catalogue, weights = w)
table(scored$score_flag)
head(scored[order(-scored$gfs_points),
            c("question_id", "std_type", "gfs_points", "score_flag", "score_basis")])
```

`score_flag` has four values:

- `auto` -- the GfS mapping is deterministic from the structure.
- `inferred` -- the structure is clear, but choosing the Table 1 category
  involved an explicit judgement (a dropdown mapped to "rating" as the
  closest category, a numeric-looking dropdown mapped to "simple
  numerical answer", a slider). The number is still determinate.
- `manual` -- a human should check; the mapping is genuinely uncertain.
- `unknown` -- an unmapped type, left unscored as `NA`. A survey with
  unknown types produces an under-count, and `burden_report()` warns.

This is the distinction between **GfS-native** quantities (question type,
response action, number of alternatives -- all specified by the scheme)
and **package-derived** ones (estimated line counts, response-unit
structure -- inferred because the `.qsf` lacks them).

## Documented inference rules where Table 1 is silent

- **Instruction-text lines.** Table 1 counts rendered "lines" of
  instruction text. A `.qsf` stores words, not a rendered width, so lines
  are estimated as words / `words_per_line` (default 12). This is a
  pragmatic conversion heuristic, not a calibrated constant; override it
  with `burden_report(words_per_line = ...)`. See `vignette("calibration")`.
- **Dropdowns** (`MC` with a `DL` selector) have no Table 1 row. They are
  scored as a rating (2.0 for up to 5 options, 3.0 for more), or as a
  simple numerical answer (1.0) when the options are numbers and the
  question asks for a quantity or a date.
- **Sliders** are not in Table 1. They are scored as a numerical answer
  when the axis label carries units, otherwise as a rating.
- **Hidden questions.** A question hidden from the respondent by injected
  CSS or JavaScript scores 0.
- **Loop & Merge blocks** are scored per iteration and multiplied by the
  block's explicit iteration cap. Within-loop display logic is not
  enumerated.

### Question stems are not separately scored

Table 1's "question or transition, up to 3 lines = 2.0" is applied to
standalone instruction blocks only. It is not added on top of every
question's response score. The GfS scheme may intend the reading cost of
a question stem to be absorbed into the item weight; settling that needs
the underlying GfS methodology, not Table 1 alone. A question with a long
disambiguation stem is therefore scored conservatively, and its stem
length is reported separately as a readability diagnostic (see
`vignette("reading-the-report")`).

### Multi-answer matrices have no Table 1 rule

A grid where each cell allows multiple answers -- "tick all that apply"
across two axes -- is scored as `rows x cols x 0.5`. Each cell is one
trivial closed yes/no decision; GfS scores a closed yes/no at 1.0,
discounted here for the working-memory efficiency of answering in a grid.
This proxy is **orientation-invariant**: symmetric in rows and columns,
so unlike an earlier `rows x 4.0` proxy it does not change with how
Qualtrics happened to store the grid. It is a documented inference, not
GfS. The weight is `gfs_weights()$matrix_cell_multi`.

## Overriding the weights

Every weight in `gfs_weights()` can be changed and passed back in:

```{r override}
w2 <- gfs_weights()
w2$rating_small <- 2.5
score_burden(catalogue, weights = w2)$gfs_points[1:5]
```

## References

- Heimgartner, D. and Axhausen, K. W. (2024). Predicting response rates
  once again. *Findings*.
  <https://findingspress.org/article/125481-predicting-response-rates-once-again>
