---
title: "Paths and display logic"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Paths and display logic}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options:
  markdown:
    wrap: 72
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

This article covers how surveyBurden reconstructs the routes through a
survey and turns them into a burden spread, and the difference between
the structural profile and a population-weighted average. For per-question
scoring see `vignette("gfs-scoring")`; for the printed output see
`vignette("reading-the-report")`.

```{r setup}
library(surveyBurden)
demo <- system.file("extdata", "demo_travel_survey.qsf", package = "surveyBurden")
qsf <- read_qsf(demo)
report <- burden_report(demo, quiet = TRUE)
```

## Survey flow

Contemporary web surveys are not a fixed sequence. `resolve_flow()` walks
the `SurveyFlow`, forking at every `Branch` into a condition-true and a
condition-false continuation, and at every `EndSurvey` into a screen-out.
Branch conditions are **not evaluated** -- they depend on embedded data
or prior answers that are unknown before fielding -- so both outcomes are
always kept. The result is every block sequence a respondent could
encounter.

```{r flow}
resolve_flow(qsf)
```

```{r flow-counts}
nrow(report$paths)
sum(report$paths$status == "complete")
sum(report$paths$status == "screen_out")
```

Block randomisers are treated as "all sub-blocks shown, in survey order"
-- the randomised subsets are not enumerated. The path space is
enumerated up to `max_paths` (10000); a flow that would produce more
raises an error rather than a partial answer.

## Display logic

Within a path, `resolve_paths()` partitions the questions into three
sets: **always shown**, **may be shown** (gated by a condition that
cannot be evaluated before fielding), and **never shown on this path**
(the gate's trigger question is not on the path).

```{r paths}
rp <- resolve_paths(qsf)
data.frame(
  path_id    = rp$path_id,
  screen_out = rp$terminates_early,
  n_always   = lengths(rp$q_always),
  n_maybe    = lengths(rp$q_maybe),
  n_gates    = rp$n_gates
)
```

Qualtrics display logic is stored as `If` / `ElseIf` / `AndIf` groups.
The package folds each group left to right: `ElseIf` clauses combined
with OR, `AndIf` clauses with AND, a missing type falling back to AND.
Flattening everything to AND would under-count the cases where a question
appears if any one of several conditions holds.

## Path enumeration

`path_burden_profile()` goes inside the "may be shown" set and enumerates
the feasible display-logic states.

- Conditional questions are grouped into **coupling components**: two
  questions couple if they share a gate, directly or through a chain of
  shared gates.
- A component with at most 5000 joint gate-states is enumerated
  **exactly** -- every joint state of its gates, respecting single-choice
  mutual exclusion and AND-across-gates conditions.
- A larger component (dozens of questions keyed off one popular trigger
  such as employment status) falls back to a **primary-gate
  approximation**: each question is scored against its first gate, the
  rest held permissive.
- Separate components are convolved as independent.
- Loop & Merge blocks are unrolled to the explicit `?v=N` cap, with the
  iteration count treated as uniform over 1 to N in the structural
  profile.

```{r profile}
pbp <- path_burden_profile(qsf)
pbp[, c("path_id", "terminates_early", "burden_min",
        "burden_median", "burden_max")]
nrow(pbp$profile[[3]])   # feasible burden values on complete path 3
```

Each row of a `profile` table is one feasible burden value with a
structural weight from the gate combinatorics and the uniform
loop-iteration assumption. **That weight is not a respondent
probability.**

## Structural versus population-weighted burden

This distinction is central and is easy to state wrongly.

The default result is a **structural burden profile**. It lists every
feasible combination of optional questions once. It carries no respondent
probabilities. When the report says the median is
`r report$burden$points[report$burden$statistic == "median"]` points, it
means the median across those feasible combinations. It does **not** mean
half of respondents face at least that much burden. Nobody has told the
package how common each combination is.

To get a real average over respondents you need observed routes. Give
`burden_report()` a data frame of respondents and it adds a
**population-weighted respondent burden**: the sum, over the observed
routes, of each route's burden multiplied by how often that route
occurred. Writing the observed frequency of route k as w_k and its
burden as B_k:

$$\bar{B} = \sum_k w_k \, B_k$$

The structural profile is what you have without the w_k. The
population-weighted result is what the w_k buy you.

The recognised route columns are `loop_<id>` (the iteration count for a
Loop & Merge block, keyed by its driving question id or block id) and
`visit_<block_id>` (`TRUE` to include a branch-gated optional block). The
demo has two loop blocks, `BL6` and `BL8`. A small constructed set of 50
respondent routes:

```{r population}
set.seed(1)
routes <- data.frame(
  loop_BL6 = rbinom(50, 5, 0.4),   # "Other adults" loop, cap 5
  loop_BL8 = rbinom(50, 3, 0.5)    # "Vehicle details" loop, cap 3
)
burden_report(demo, routes = routes, quiet = TRUE)$population
```

Real route data would come from a fielded response file, not from
`rbinom()`. When `routes` are supplied the report's verdict line also
gives the population-weighted median.

Without routes, `respondent_burden()` still gives a per-path table and
`summary_line()` turns it into a sentence:

```{r respondent}
summary_line(respondent_burden(demo))
```
