---
title: "Reading the burden report"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Reading the burden report}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options:
  markdown:
    wrap: 72
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

This article walks the printed `burden_report()` output section by
section: what each shows, how to read it, and when it is telling you
something is wrong. For how the underlying numbers are produced see
`vignette("gfs-scoring")` and `vignette("paths-and-display-logic")`.

```{r setup}
library(surveyBurden)
demo <- system.file("extdata", "demo_travel_survey.qsf", package = "surveyBurden")
report <- burden_report(demo, quiet = TRUE)
report
```

## The verdict line

The first line under the title is a plain-language summary: the median
completing-path burden in GfS points and in minutes, its ratio to the
399-point GfS benchmark, and the point range across the survey's
completing paths. When you supply `routes` it also reports the
population-weighted median (see
`vignette("paths-and-display-logic")`).

`summary(report)` prints the same verdict with the structural counts and
nothing else.

## Instrument

```{r instrument}
report$instrument
```

Structural counts read straight off the survey definition: live
questions, blocks, branch points, block randomisers, Loop & Merge
blocks, and early-exit (`EndSurvey`) points. These describe the
programmed instrument, not any respondent. A high branch or early-exit
count means the survey routes people quite differently, so a single
burden figure would describe few of them.

## Paths

```{r paths}
report$paths[, c("path_id", "status", "n_blocks",
                 "burden_floor", "burden_ceiling")]
```

Each **structural path** is one distinct block sequence a respondent
could take. `status` is `complete` (reaches a real submission) or
`screen_out` (hits an `EndSurvey` first). `burden_floor` and
`burden_ceiling` are the naive per-path bounds -- only always-shown
questions, loops run once, versus every reachable question with loops at
their cap.

The printed "Display-logic combinations checked" line is the number of
feasible skip-logic states enumerated per complete path.

## Burden

```{r burden}
report$burden
attr(report$burden, "basis")
```

- `points` is the GfS point total; `minutes` is `points` divided by the
  points-per-minute rate (default 12); `index` is `points / 1500`, a
  0--1 rescaling where 1.0 is the level Heimgartner and Axhausen's sample
  found rare.
- `attr(, "basis")` is `"structural"` when the display-logic profile was
  enumerated (the default), `"naive"` when `profile = FALSE`, and
  `"none"` when no path completes. The naive band scores each optional
  question as independently hideable for the minimum and independently
  shown for the maximum; that is faster, but its minimum can fall below
  anything actually reachable and it reports no median. **Quote the
  structural basis.**
- The **benchmark** line is the published GfS reference point: a median
  of 399 points across the 79 scored survey waves in Heimgartner and
  Axhausen (2024). The verdict line's ratio is the survey's median over
  that 399.

The median here is the middle value across the feasible skip-logic
combinations, **not** the burden half of respondents exceed --
`vignette("paths-and-display-logic")` returns to this.

Per-path structural figures are in `$paths`:

```{r paths-burden}
report$paths[, c("path_id", "status", "burden_min",
                 "burden_median", "burden_max")]
```

These columns are `NA` when `profile = FALSE`.

## Burden by block

```{r blocks}
report$blocks[, c("block_name", "n_questions", "gfs_points", "share")]
```

Question burden totalled by block, in survey order, with each block's
share of the all-questions-shown total. The printed bar chart is this
`share` column. Use it to see which modules drive the length.

## Highest-burden questions

The printed table lists the six questions with the largest single
`gfs_points`. The full per-question detail is `report$items`, including
the `score_basis` string that records the rule used for each:

```{r items}
head(report$items[order(-report$items$gfs_points),
                  c("question_id", "std_type", "gfs_points",
                    "score_flag", "score_basis")], 3)
```

`score_flag` is `auto` (deterministic mapping), `inferred` (the mapping
needed a documented judgement), `manual` or `unknown`. See
`vignette("gfs-scoring")`.

## Readability diagnostics

```{r readability}
report$readability$stem_threshold
nrow(report$readability$long_stems)
nrow(report$readability$long_labels)
report$readability$long_grids[, c("question_id", "block_name", "n_rows")]
```

Three counts: question stems over `stem_threshold` words (default 40),
matrix row/option labels over `label_threshold` words (default 10), and
matrix questions with more than 6 rows. These are **reading-load
signals** reported separately -- **none of them changes any GfS score**.
A long grid or a wordy stem is flagged so it is visible without being
silently priced into the burden total. Long grids in particular invite
satisficing (answering the same way down every row).

## Calculation certainty

```{r certainty}
report$certainty$paths[c("n_full", "n_exact", "n_with_unresolved")]
report$certainty$display_logic[c("n_conditional", "n_exact", "n_approx")]
report$certainty$loops[c("n_known", "n_unknown")]
report$certainty$scores
```

For each survey this states which parts of the calculation are exact and
which rest on a documented approximation: how many paths resolve to a
single burden value, how many conditional questions were enumerated
exactly versus approximated, whether every loop has a known cap, and the
`auto` / `inferred` / `manual` / `unknown` split of the item scores. A
result can be fully reproducible without every underlying interpretation
being certain; this block makes the uncertain parts countable. Pass
`certainty = FALSE` to skip it.

## QC warnings

The printed report ends with plain-language flags: unscored question
types (the reported burden is then an under-count), long stems, long
labels, long grids, a high screen-out share, and a reminder that the
structural range is not a respondent probability. The full vector is
`report$warnings`.

```{r warnings}
report$warnings[1]
```

## References

- Heimgartner, D. and Axhausen, K. W. (2024). Predicting response rates
  once again. *Findings*.
  <https://findingspress.org/article/125481-predicting-response-rates-once-again>
