---
title: "Get started with surveyBurden"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Get started with surveyBurden}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options:
  markdown:
    wrap: 72
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

`surveyBurden` estimates **ex-ante instrument burden**: the response
effort a programmed Qualtrics survey demands, worked out before fielding
from the instrument alone. It reads the survey, scores every question
with the GfS+ point scheme (extending Heimgartner and Axhausen 2024), reconstructs the paths
the flow and display logic allow, and reports how burden varies across
those paths.

In the terms of Yan and Williams (2022) this is the *instrument* side of
respondent burden -- length, difficulty, and the effort the design
requires. It is not the *perceived* burden a particular respondent
reports feeling, and it is not a completion-time predictor.

This article gets you to a first report. Which function you need depends
on what you have:

| What you have | Start with | What you get | Guide |
|---|---|---|---|
| A Qualtrics survey (`.qsf` or API) | `burden_report(qsf)` | a model-based structural burden profile across the survey's paths | `vignette("reading-the-report")` |
| That, plus respondent routes and loop counts | `burden_report(qsf, routes = routes)` | a distribution of respondent-level predicted burden | `vignette("paths-and-display-logic")` |
| A response export | `realised_burden(qsf, responses)` | answer-based burden per respondent | `vignette("realised-burden")` |
| Completion times | `validate_times(qsf, observed)` | a check of predicted against observed time, and a survey-specific rate | `vignette("calibration")` |
| A draft you want to make lighter | `burden_report()` before and after an edit | where the burden is, why, and how much a change removes | `vignette("survey-revision")` |
| A survey from another platform | `score_burden(validate_catalogue(catalogue))` | per-question scores and a total | `vignette("extensions")` |

The other articles go deeper:

- `vignette("reading-the-report")` -- what each part of the printed
  output means.
- `vignette("survey-revision")` -- finding the heaviest parts of a draft,
  changing them, and comparing before and after.
- `vignette("gfs-scoring")` -- how a Qualtrics widget becomes points, and
  what the automatic scorer does and does not apply.
- `vignette("paths-and-display-logic")` -- flow, skip logic, the
  structural-versus-population-weighted distinction, and building
  respondent routes.
- `vignette("calibration")` -- the `words_per_line` and points-per-minute
  assumptions, validating against completion times, and the limitations.
- `vignette("realised-burden")` -- answer-based burden from response
  data.
- `vignette("extensions")` -- what works for surveys without a QSF, and
  what full support for another platform would take.

## Install

```r
# install.packages("pak")
pak::pak("pmpk20/surveyBurden")
```

## One call

```{r setup}
library(surveyBurden)
demo <- system.file("extdata", "demo_travel_survey.qsf", package = "surveyBurden")
```

`burden_report()` runs the whole pipeline. Every code chunk in these
articles runs against the small fictional "Neighbourhood Travel Survey"
that ships with the package, so the numbers shown are the numbers the
current package produces. To use your own survey, replace `demo` with a
path to a `.qsf` file, a bare `SV_...` survey id, or a Qualtrics
survey-builder URL (see `?fetch_qsf` for fetching through the
Qualtrics API).

```{r quick-start}
report <- burden_report(demo, quiet = TRUE)
report
```

The printout opens with a one-line **verdict**: the median
completing-path burden in points and minutes, its ratio to the 399-point
GfS benchmark, and the range across completing paths. Below it:

- **Instrument** -- structural counts (questions, blocks, branch points,
  loop blocks, early exits).
- **Paths** -- how many distinct paths the flow allows, and how many
  complete rather than screen out.
- **Burden** -- the headline spread of path burden, in points, minutes
  and an index (points / 1500, so 1 is a rare burden), with the published benchmark underneath.
- **Burden by block**, **Highest-burden questions**, **Readability
  diagnostics**, **Calculation certainty**, **QC warnings**.

`vignette("reading-the-report")` walks each of these in turn.

## Vocabulary

These terms are used throughout, with fixed meanings. The key distinction: a **path** is a possibility in the survey's design; a **route** is what one respondent actually did.

| term | meaning |
|---|---|
| **path** | one way through the survey that its flow allows: the blocks shown, given which branches are taken. Worked out from the survey design alone; no respondent is involved. |
| **route** | the way one respondent actually went: the optional blocks they entered and how many times they looped. Comes from response data (the `routes` argument). Many respondents' routes can fall on the same path. |
| **question burden** | the GfS+ point score for one question, from its type and structure |
| **path burden** | the total question burden along one path. Display logic and loops can vary within a path, so a path has a spread of burdens rather than one number. |
| **structural burden profile** | the burden values a path's modelled display-logic states produce, with model weights (equal per state; each path's weights sum to 1). Not a respondent probability distribution. |
| **predicted respondent burden** | one respondent's predicted burden: the burden of the path matching their route, with their own loop counts (`pred_pts`). |
| **population-weighted respondent burden** | predicted respondent burden summarised across respondents, so each path counts as often as respondents took it. Produced only when `routes` are supplied. |
| **realised burden** | the burden of the questions a respondent actually answered, from response data (`realised_burden()`). |
| **calculation certainty** | counts, for a given survey, of the parts of the calculation that are enumerated in full within the model and the parts that rest on documented approximations. A modelling diagnostic, not a check against the live survey. |
| **readability diagnostics** | counts of long question stems, long matrix labels, and long grids. Reading-load signals, reported separately, never folded into the GfS+ score. |

## The report object

`report` is a structured object, not printed text. Pull out each part:

```{r components}
report$burden      # min / p25 / median / p75 / max, in points, minutes and index
head(report$items[order(-report$items$gfs_points),
                  c("question_id", "std_type", "gfs_points", "score_flag")])
```

Other components: `report$instrument` (the structural counts),
`report$paths` (one row per structural path), `report$blocks` (burden by
block), `report$readability`, `report$certainty` and `report$warnings`.
Scalars such as the points-per-minute rate live on attributes:
`attr(report, "points_per_minute")`.

`summary(report)` prints just the headline:

```{r summary}
summary(report)
```

## A short worked example

```{r worked}
report$paths[, c("path_id", "status", "n_questions_floor",
                 "n_questions_ceiling", "burden_min", "burden_max")]
```

The two screen-out paths end early after the area check; the two complete
paths differ mainly in one branch-gated block. The spread from minimum to
maximum is the effect of the survey's optional questions and loop
iterations, not measurement noise.

## References

- Heimgartner, D. and Axhausen, K. W. (2024). Predicting response rates
  once again. *Findings*.
  <https://findingspress.org/article/125481-predicting-response-rates-once-again>
- Yan, T. and Williams, D. (2022). Response burden: a review and
  conceptual framework. *Journal of Official Statistics*.
