---
title: "Revising a survey to reduce burden"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Revising a survey to reduce burden}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options:
  markdown:
    wrap: 72
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

**What this answers.** Which part of a draft survey carries the most
burden, why it scores as it does, and how much a specific change reduces
the modelled burden.

**What you need.** The draft's `.qsf` (export it from Qualtrics: *Tools >
Import/Export > Export survey*), and the revised version's `.qsf` after
you edit it.

```{r setup}
library(surveyBurden)
demo <- system.file("extdata", "demo_travel_survey.qsf", package = "surveyBurden")
before <- burden_report(demo, quiet = TRUE)
```

## 1. Find where the burden is

Start with blocks, then the questions inside the heaviest ones:

```{r blocks}
b <- before$blocks
b[order(-b$share), c("block_name", "n_questions", "gfs_points", "share")][1:3, ]

it <- before$items
it[order(-it$gfs_points),
   c("question_id", "block_name", "std_type", "n_rows", "n_cols", "gfs_points")][1:4, ]
```

A block's share is of the points of all its questions, whether or not a
given path shows them; `before$paths` shows which paths include which
blocks.

## 2. Check how the heavy questions were scored

Before changing a question, check that its score reflects the question
rather than an inference rule you disagree with:

```{r basis}
it[it$question_id %in% c("QID21", "QID19"), c("question_id", "score_flag", "score_basis")]
```

Both are `auto`: a single-answer matrix scores one rating per row, and a
rating with more than 5 options costs more than one with up to 5 (see
`vignette("gfs-scoring")`). So `QID21`, a 6-statement agreement grid on a
7-point scale, costs 6 x 3 = 18 points. Two levers follow from the rule:
fewer statements, or a 5-point scale.

If a heavy question is `inferred`, read its `score_basis` first: the
burden may come from a mapping (for example, a slider scored as a
numerical answer) rather than from the question's design.

## 3. Make the change and re-score

Make the edit in Qualtrics and export the revised `.qsf`. To keep this
article self-contained, the same edit is made here in R: keep 4 of the 6
statements and 5 of the 7 scale points.

```{r revise}
revised <- read_qsf(demo)
i <- which(vapply(revised$SurveyElements, function(e)
  identical(e$Element, "SQ") && identical(e$Payload$QuestionID, "QID21"),
  logical(1)))
p <- revised$SurveyElements[[i]]$Payload
p$Choices <- p$Choices[1:4]   # statements (rows)
p$Answers <- p$Answers[1:5]   # scale points (columns)
revised$SurveyElements[[i]]$Payload <- p

after <- burden_report(revised, quiet = TRUE)
```

With a real revised export you would simply run
`after <- burden_report("my_survey_v2.qsf")`.

## 4. Compare

```{r compare}
data.frame(
  statistic = before$burden$statistic,
  before    = before$burden$points,
  after     = after$burden$points,
  change    = after$burden$points - before$burden$points
)

q <- "QID21"
rbind(before = before$items[before$items$question_id == q, c("n_rows", "n_cols", "gfs_points")],
      after  = after$items[after$items$question_id == q, c("n_rows", "n_cols", "gfs_points")])
```

`QID21` falls from 18 to 8 points. It is shown on every path, so every
quantile falls by the same 10 points. A change to a question that only
some paths show would move the quantiles by different amounts; compare
`before$paths` and `after$paths` to see which paths changed.

To see which lever did the work, change one thing at a time: here,
cutting to 4 rows at 7 points would give 4 x 3 = 12, and keeping 6 rows
at 5 points would give 6 x 2 = 12.

## What the comparison establishes

It establishes that, under the GfS+ scheme and the package's model of the
survey's routing, the revision lowers modelled burden by the amount shown,
and which questions and paths the change comes from.

It does **not** establish:

- that completion time falls by `change / 12` minutes -- the
  points-to-minutes rate is a rule of thumb (see `vignette("calibration")`);
- that data quality, response rates or dropout improve -- burden scores
  are not predictions of respondent behaviour;
- that the shorter question measures the same thing -- dropping
  statements or scale points is a measurement decision, not a burden one.

## Next

- Fielded the revision? Compare answer-based burden with
  `realised_burden()` (`vignette("realised-burden")`), and check predicted
  against observed completion times with `validate_times()`
  (`vignette("calibration")`).
- Unsure why a question scored as it did? `vignette("gfs-scoring")` lists
  the rule for every question type.
