---
title: "Get started with surveyBurden"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Get started with surveyBurden}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options:
  markdown:
    wrap: 72
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

`surveyBurden` estimates **ex-ante instrument burden**: the response
effort a programmed Qualtrics survey demands, worked out before fielding
from the instrument alone. It reads the survey, scores every question
with the published GfS / Axhausen point scheme, reconstructs the routes
the flow and display logic allow, and reports how burden varies across
those routes.

In the terms of Yan and Williams (2022) this is the *instrument* side of
respondent burden -- length, difficulty, and the effort the design
requires. It is not the *perceived* burden a particular respondent
reports feeling, and it is not a completion-time predictor.

This article gets you to a first report. The others go deeper:

- `vignette("reading-the-report")` -- what each part of the printed
  output means.
- `vignette("gfs-scoring")` -- how a Qualtrics widget becomes points.
- `vignette("paths-and-display-logic")` -- flow, skip logic, and the
  structural-versus-population-weighted distinction.
- `vignette("calibration")` -- the `words_per_line` and points-per-minute
  assumptions, and the limitations.

## Install

```r
# install.packages("pak")
pak::pak("pmpk20/surveyBurden")
```

## One call

```{r setup}
library(surveyBurden)
demo <- system.file("extdata", "demo_travel_survey.qsf", package = "surveyBurden")
```

`burden_report()` runs the whole pipeline. Every code chunk in these
articles runs against the small fictional "Neighbourhood Travel Survey"
that ships with the package, so the numbers shown are the numbers the
current package produces. To use your own survey, replace `demo` with a
path to a `.qsf` file, a bare `SV_...` survey id, or a Qualtrics
survey-builder URL (see `vignette("paths-and-display-logic")` for the API
route).

```{r quick-start}
report <- burden_report(demo, quiet = TRUE)
report
```

The printout opens with a one-line **verdict**: the median
completing-path burden in points and minutes, its ratio to the 399-point
GfS benchmark, and the range across completing paths. Below it:

- **Instrument** -- structural counts (questions, blocks, branch points,
  loop blocks, early exits).
- **Paths** -- how many distinct routes the flow allows, and how many
  complete rather than screen out.
- **Burden** -- the headline spread of path burden, in points, minutes
  and a 0--1 index, with the published benchmark underneath.
- **Burden by block**, **Highest-burden questions**, **Readability
  diagnostics**, **Calculation certainty**, **QC warnings**.

`vignette("reading-the-report")` walks each of these in turn.

## Vocabulary

Five terms are used throughout, with fixed meanings:

| term | meaning |
|---|---|
| **question burden** | the GfS point score for one question, from its type and structure |
| **path burden** | the total question burden along one structural path through the survey flow |
| **structural burden profile** | the set of feasible burden values across a path's display-logic sub-states, each counted once. Not a probability distribution. |
| **population-weighted respondent burden** | burden averaged over real respondent routes, where route frequencies stand in for the sub-state probabilities. Produced only when `routes` are supplied. |
| **calculation certainty** | the split, for a given survey, between parts of the calculation that are exact and parts that rest on documented approximations |
| **readability diagnostics** | counts of long question stems, long matrix labels, and long grids. Reading-load signals, reported separately, never folded into the GfS score. |

## The report object

`report` is a structured object, not printed text. Pull out each part:

```{r components}
report$burden      # min / p25 / median / p75 / max, in points, minutes and index
head(report$items[order(-report$items$gfs_points),
                  c("question_id", "std_type", "gfs_points", "score_flag")])
```

Other components: `report$instrument` (the structural counts),
`report$paths` (one row per structural path), `report$blocks` (burden by
block), `report$readability`, `report$certainty` and `report$warnings`.
Scalars such as the points-per-minute rate live on attributes:
`attr(report, "points_per_minute")`.

`summary(report)` prints just the headline:

```{r summary}
summary(report)
```

## A short worked example

```{r worked}
report$paths[, c("path_id", "status", "n_questions_floor",
                 "n_questions_ceiling", "burden_min", "burden_max")]
```

The two screen-out paths end early after the area check; the two complete
paths differ mainly in one branch-gated block. The spread from minimum to
maximum is the effect of the survey's optional questions and loop
iterations, not measurement noise.

## References

- Heimgartner, D. and Axhausen, K. W. (2024). Predicting response rates
  once again. *Findings*.
  <https://findingspress.org/article/125481-predicting-response-rates-once-again>
- Yan, T. and Williams, D. (2022). Response burden: a review and
  conceptual framework. *Journal of Official Statistics*.
